The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-11-05 script/en/chapters/cryptomethods.tex
CT script: Many small changes; especially referenced CT2, JCT and Sage where previously only CT1 was mentioned.

% $Id$
% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung / Syntax rules English:
%         - Chapter header with all nouns/verbs-words capitalized.
%         - All other headers in the normal lower/upper case manner
%           like sentences, but without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{~\csname @arabic\endcsname \c@page~} \makeatother
\hypertarget{Kapitel_1}{}
\chapter{Security Definitions and Encryption Procedures}
\label{Label_Kapitel_1}
(Bernhard Esslinger, Joerg-Cornelius Schneider, May 1999; Updates Dec. 2001, Feb. 2003,
June 2005, July 2007, January 2010, June 2012)

\begin{center}
\fbox{\parbox{15cm}{%
{\em Saying from India}:\\
Explain it to me, I will forget it.\\
Show it to me, maybe I will remember it.\\
Let me do it, and I will be good at it.}}
\end{center}

@@

@@

@@

% --------------------------------------------------------------------------
\section[Symmetric encryption]
{Symmetric encryption\footnotemark}
  \footnotetext{%
    With {\bf CrypTool 1}\index{CrypTool 1} you can execute the following modern
    symmetric encryption algorithms 
    (using the menu path {\bf Crypt \textbackslash{} Symmetric (modern)}):\\
    IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
    MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
    Twofish (AES candidate), Rijndael (official AES algorithm)\index{AES}.

    With {\bf CrypTool 2}\index{CrypTool 2} you can execute the following modern
    symmetric encryption algorithms (using in the Startcenter {\bf Templates
    \textbackslash{} Cryptography \textbackslash{} Modern \textbackslash{}
    Symmetric}):\\
    AES, DES, PRESENT, RC2, RC4, SDES, TEA, Triple-DES, Twofish.

    In {\bf JCrypTool}\index{JCrypTool} you can execute the following modern
    symmetric encryption algorithms:\\ 
    AES, Rijndael, Camellia, DES, Dragon, IDEA, LFSR, MARS, Misty1, RC2, RC5,
    RC6, SAFER+, SAFER++, Serpent, Shacal, Shacal2, Twofish.
  }
\nopagebreak
For {\em symmetric} encryption \index{Encryption!symmetric} sender and
recipient must be in possession of a common (secret) key which they have
exchanged before actually starting to communicate. The sender uses this
key to encrypt the message and the recipient uses it to decrypt it.
\par \vskip + 3pt

@@

The most well-known symmetric encryption procedure is the \index{DES} DES algorithm.
The DES algorithm has been developed by IBM in collaboration with the
National Security Agency \index{NSA} (NSA), and was published as a standard in
1975. Despite the fact that the procedure is relatively old, no effective attack
on it has yet been detected. The most effective way of attacking consists of
testing (almost) all possible keys until the right one is found ({\em brute-force-attack}).
\index{Attack!brute-force} Due to the relatively short key length of
effectively 56 bits (64 bits, which however include 8 parity bits), numerous
messages encrypted using DES have in the past been broken. Therefore, the
procedure can now only be considered to be conditionally secure. Symmetric
alternatives to the DES procedure include the IDEA \index{IDEA} or Triple-DES
algorithms.\par \vskip + 3pt

@@

More details about the AES algorithms and the AES candidates of the last round
can be found i.e. within the online help of CrypTool 1\index{CrypTool 1}%
\footnote{%
      CrypTool 1 online help\index{CrypTool 1}: The index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }.

@@

% --------------------------------------------------------------------------
\subsection{Current status of brute-force attacks on symmetric algorithms (RC5)}
\index{Attack!brute-force}
\index{RC5}
\label{Brute-force-gegen-Symmetr}

@@

Brute-force (exhaustive search, trial-and-error) means to completely examine all keys of the key space: so no special analysis methods have to be used. Instead, the ciphertext is decrypted with all possible keys and for each resulting text it is checked, whether this is a meaningful clear text. A key length of 64 bit means at most $2^{64}$ = 18,446,744,073,709,551,616 or about 18 trillion (GB) / 18 quintillion (US)  keys to check\index{CrypTool}%
\footnote{%
    With CrypTool\index{CrypTool} you can also try brute-force attacks
    of modern symmetric algorithms (using the menu path
    {\bf Analysis \textbackslash{} Symmetric Encryption (modern)}): here
    the weakest knowledge of an attacker is assumed, he performs a 
    ciphertext-only attack.
    To achieve a result in an appropriate time with a single PC you should 
    mark not more than 20 bit of the key as unknown.
}.

Companies like RSA Security provide so-called cipher challenges in order
to quantify the security offered by well-known symmetric ciphers as DES,
Triple-DES or RC5%
\index{DES}\index{RC5}\index{Triple-DES}%
\footnote{%
 \url{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}\\
 In May 2007 RSA announced that they will not confirm the correctness of
 the not yet solved RC5-72 challenge.}.
They offer prizes for those who manage to decipher ciphertexts, encrypted with different algorithms and different key lengths, and to unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

@@

There are also cipher challenges\footnote{%
  A wide spectrum of simple and complex crypto riddles are included
  in the ``MysteryTwister C3''~challenge: \url{http://www.mysterytwisterc3.org}.
  \index{Challenge}
  \index{Cipher challenge}
  \index{Crypto challenge}
  }  for asymmetric algorithms (please see chapter \ref{nt:NoteFactorization}).

% --------------------------------------------------------------------------
\section[Asymmetric encryption]
{Asymmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool\index{CrypTool} you can execute RSA encryption
    and decryption (using the menu path {\bf Crypt \textbackslash{} Asymmetric}).
    In both cases you must select a RSA key pair. Only in the case of decryption
    the secret RSA key is necessary: so here you are asked to enter the PIN. 
  }
\nopagebreak
In the case of {\em asymmetric} encryption \index{Encryption!asymmetric} each
subscriber has a personal pair of keys consisting of a {\em secret}
\index{Key!secret} key and a {\em public} key\index{Key!public}. The public
key, as its name implies, is made public, e.g. in a key directory on the
Internet.\par \vskip + 3pt

@@

The most well-known asymmetric procedure is the \index{RSA} 
RSA algorithm\index{CrypTool}%
\footnote{%
The RSA algorithm is extensively described in chapter \ref{rsabeweis} and later
within this script. 
The RSA cryptosystem can be executed in many variations with 
CrypTool\index{CrypTool} (using the menu path
{\bf Individual Procedures \textbackslash{} RSA Cryptosystem \textbackslash{}
RSA Demonstration}). 
The topical research results concerning RSA are described 
in chapter \ref{SecurityRSA}.
}%
, named after its developers Ronald \index{Rivest, Ronald} Rivest, Adi
\index{Shamir, Adi} Shamir and Leonard \index{Adleman, Leonard} Adleman. The RSA algorithm
was published in 1978. The concept of asymmetric encryption was first
introduced by Whitfield Diffie \index{Diffie, Whitfield}  and Martin
\index{Hellman, Martin} Hellman in 1976. Today, the ElGamal \index{ElGamal, Tahir}
procedures also play a decisive role, particularly the \index{Schnorr, C.P.} Schnorr
variant in the \index{DSA} DSA (Digital \index{Signature!digital}Signature
Algorithm).

% --------------------------------------------------------------------------
% \newpage
\section[Hybrid procedures]{Hybrid procedures\footnotemark}
\footnotetext{%
Within CrypTool\index{CrypTool} you can find this technique using the menu
path {\bf Crypt \textbackslash{} Hybrid}: 
There you can follow the single steps and its dependencies with concrete
numbers. The variant with RSA as the asymmetric algorithm is graphically
visualized; the variant with ECC uses the standard dialogs. In both cases
AES is used as the symmetric algorithm.\index{AES}
}
\label{CM_Hybrid-procedures}
\index{Hybrid procedure}

@@

In this case the bulk data is encrypted using symmetric procedures: The key
used for this is a secret session key\index{Session key} generated by the
sender randomly\footnote{%
An important part of cryptographically secure techniques is to generate 
random numbers. Within CrypTool\index{CrypTool} you can check out
different random number generators using the menu path
{\bf Indiv. Procedures \textbackslash{} Generate Random Numbers}. 
Using the menu path {\bf Analysis \textbackslash{} Analyze Randomness}
you can apply different test methods for random data to binary documents. \\
Up to now CrypTool has concentrated on cryptographically strong 
pseudo random number generators. Only the integrated Secude\index{SECUDE IT Security}
generator involves a "pure" random source. 
}\index{Random}
that is only used for this message.

@@

% --------------------------------------------------------------------------
\newpage
% \vskip +40 pt
\begin{center}
\fbox{\parbox{15cm}{
    \emph{IETF\footnotemark:} \\
    There is an old saying inside the US National Security Agency (NSA):\\
    "Attacks always get better; they never get worse."
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }
\section{Cryptanalysis and ciphers for educational purposes} \index{Cryptanalysis}

@@

@@

>>>>

% $Id$
% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung / Syntax rules English:
%         - Chapter header with all nouns/verbs capitalized.
%         - All other headers in the normal lower/upper case manner
%           like sentences, but without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{~\csname @arabic\endcsname \c@page~} \makeatother
\hypertarget{Kapitel_1}{}
\chapter{Security Definitions and Encryption Procedures}
\label{Label_Kapitel_1}
(Bernhard Esslinger, Joerg-Cornelius Schneider, May 1999; Updates Dec. 2001, Feb. 2003,
June 2005, July 2007, Jan. 2010, Nov. 2012)

@@

For further explanations we use the notation from the following Figure \ref{Generic-Notations-when-Encrypting}:
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.7]{figures/Generic-Notation-Encryption_en.png}
\caption{Common notations when using ciphers} 
\label{Generic-Notations-when-Encrypting}
\end{center}
\end{figure}

@@

\begin{ctsquote}
\caption{Saying from India}
Explain it to me, I will forget it.\\
Show it to me, maybe I will remember it.\\
Let me do it, and I will be good at it.
\end{ctsquote}

@@

% --------------------------------------------------------------------------
\section[Symmetric encryption]
{Symmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool 1 ({\bf CT1})\index{CrypTool 1} you can execute the following modern
    symmetric encryption algorithms 
    (using the menu path {\bf Crypt \textbackslash{} Symmetric (modern)}):\\
    IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
    MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
    Twofish (AES candidate), Rijndael (official AES algorithm)\index{AES}.\\
    With CrypTool 2 ({\bf CT2})\index{CrypTool 2} you can execute the following modern
    symmetric encryption algorithms (using in the Startcenter {\bf Templates
    \textbackslash{} Cryptography \textbackslash{} Modern \textbackslash{}
    Symmetric}):\\
    AES, DES, PRESENT, RC2, RC4, SDES, TEA, Triple-DES, Twofish.\\
    In JCrypTool ({\bf JCT})\index{JCrypTool} you can execute the following modern
    symmetric encryption algorithms:\\ 
    AES, Rijndael, Camellia, DES, Dragon, IDEA, LFSR, MARS, Misty1, RC2, RC5,
    RC6, SAFER+, SAFER++, Serpent, Shacal, Shacal2, Twofish.
  }
% --------------------------------------------------------------------------
\begin{ctsquote}
``Transparency. That's the best one can hope for in a technologically advanced society ... otherwise you will just be manipulated.''
\caption[Daniel Suarez, Freedom]{Daniel Suarez\footnotemark}\index{Suarez, Daniel}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{Daniel Suarez, ``Freedom'', Dutton Adult, 2010,
                          Chapter 5, ``Getting with the Program'', p. 63, Price.}

\nopagebreak
For {\em symmetric} encryption \index{Encryption!symmetric} sender and
recipient must be in possession of a common (secret) key which they have
exchanged before actually starting to communicate. The sender uses this
key to encrypt the message and the recipient uses it to decrypt it.
\par \vskip + 3pt

@@

The most well-known symmetric encryption procedure is the \index{DES} DES algorithm.
The DES algorithm has been developed by IBM in collaboration with the
National Security Agency \index{NSA} (NSA), and was published as a standard in
1975. Despite the fact that the procedure is relatively old, no effective attack
on it has yet been detected. The most effective way of attacking consists of
testing (almost) all possible keys until the right one is found ({\em brute-force-attack}).
\index{Attack!brute-force} Due to the relatively short key length of
effectively 56 bits (64 bits, which however include 8 parity bits), numerous
messages encrypted using DES have in the past been broken. Therefore, the
procedure can not be considered secure any longer. Alternatives to the DES
procedure include IDEA\index{IDEA}, Triple-DES and especially AES.
\par \vskip + 3pt

@@

More details and further references about the AES algorithms and the AES candidates
of the last round can be found i.e. within the online help of
CrypTool 1\index{CrypTool 1}%
\footnote{%
      CrypTool 1 online help\index{CrypTool 1}: The index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}\\
      A comprehensive description of AES including C code can be found
      in \cite{cm:Haan2008}.
  }.

@@

% --------------------------------------------------------------------------
\subsection{Current status of brute-force attacks on symmetric algorithms}
\index{Attack!brute-force}
\index{RC5}
\label{Brute-force-gegen-Symmetr}

@@

Brute-force (exhaustive search, trial-and-error) means to completely examine all keys of the key space: so no special analysis methods have to be used. Instead, the ciphertext is decrypted with all possible keys and for each resulting text it is checked, whether this is a meaningful clear text. A key length of 64 bit means at most $2^{64}$ = 18,446,744,073,709,551,616 or about 18 trillion (GB) / 18 quintillion (US)  keys to check\index{CrypTool}%
\footnote{%
    With CrypTool 1\index{CrypTool 1} you can also perform brute-force attacks
    of modern symmetric algorithms (using the menu path
    {\bf Analysis \textbackslash{} Symmetric Encryption (modern)}): Here
    the weakest knowledge of an attacker is assumed, he performs a 
    ciphertext-only attack.
    To achieve a result in an appropriate time with a single PC you should 
    mark not more than 24 bit of the key as unknown.\\
    With CrypTool 2\index{CrypTool 2} you can also perform brute-force attacks
    (using the templates under {\bf Cryptanalysis \textbackslash{} Modern}).
    Highly powerful is the KeySearcher component, which can be used to distribute
    the calculations to many different computers.
}.

Companies like RSA Security provided so-called cipher challenges in order
to quantify the security offered by well-known symmetric ciphers as DES,
Triple-DES or RC5%
\index{DES}\index{RC5}\index{Triple-DES}%
\footnote{%
 \url{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}\\
 In May 2007 RSA announced that they will not confirm the correctness of
 the not yet solved RC5-72 challenge.}.
They offer prizes for those who managed to decipher ciphertexts, encrypted with different algorithms and different key lengths, and to unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

@@

There are also cipher challenges\footnote{%
  A wide spectrum of simple and complex, symmetric and asymmetric crypto riddles
  are included in the ``MysteryTwister C3''~challenge:
  \url{http://www.mysterytwisterc3.org}.
  \index{MTC3}
  \index{Challenge}
  \index{Cipher challenge}
  \index{Crypto challenge}
  }  for asymmetric algorithms (please see chapter \ref{nt:NoteFactorization}).

% --------------------------------------------------------------------------
\section[Asymmetric encryption]
{Asymmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool 1\index{CrypTool 1} you can execute RSA encryption
    and decryption (using the menu path {\bf Crypt \textbackslash{} Asymmetric}).
    In both cases you must select a RSA key pair. Only in the case of decryption
    the secret RSA key is necessary -- so here you are asked to enter the PIN.\\
    With CrypTool 2\index{CrypTool 2} you can also perform asymmetic methods
    (using the templates under {\bf Cryptography \textbackslash{} Modern}).\\
    JCrypTool\index{JCrypTool} offers asymmetic methods like RSA both within
    the {\bf Visuals} menu of the Default Perspective as well as within the
    Functional Perspective.
  }
\nopagebreak
In the case of {\em asymmetric} encryption \index{Encryption!asymmetric} each
subscriber has a personal pair of keys consisting of a {\em secret}
\index{Key!secret} key and a {\em public} key\index{Key!public}. The public
key, as its name implies, is made public, e.g. in a key directory on the
Internet.\par \vskip + 3pt

@@

The most well-known asymmetric procedure is the \index{RSA} 
RSA algorithm\index{CrypTool}%
\footnote{%
The RSA algorithm is extensively described in chapter \ref{rsabeweis} and later
within this script. 
The topical research results concerning RSA are described 
in chapter \ref{SecurityRSA}.\\
The RSA cryptosystem can be executed in many variations with 
CT1\index{CrypTool 1} (using the menu path
{\bf Individual Procedures \textbackslash{} RSA Cryptosystem \textbackslash{}
RSA Demonstration}). 
}%
, named after its developers Ronald \index{Rivest, Ronald} Rivest, Adi
\index{Shamir, Adi} Shamir and Leonard \index{Adleman, Leonard} Adleman.
The RSA algorithm
was published in 1978. The concept of asymmetric encryption was first
introduced by Whitfield Diffie \index{Diffie, Whitfield}  and Martin
\index{Hellman, Martin} Hellman in 1976. Today, the ElGamal \index{ElGamal, Tahir}
procedures also play a decisive role, particularly the \index{Schnorr, C.P.} Schnorr
variant in the \index{DSA} DSA (Digital \index{Signature!digital}Signature
Algorithm).

% --------------------------------------------------------------------------
% \newpage
\section[Hybrid procedures]{Hybrid procedures\footnotemark}
\footnotetext{%
   Within CT1\index{CrypTool 1} you can find this technique using the menu
   path {\bf Crypt \textbackslash{} Hybrid}: 
   There you can follow the single steps and its dependencies with concrete
   numbers. The variant with RSA as the asymmetric algorithm is graphically
   visualized; the variant with ECC uses the standard dialogs. In both cases
   AES is used as the symmetric algorithm.\index{AES}\\
   JCrypTool\index{JCrypTool} offers hybrid methods like ECIES within
   the Functional Perspective under {\bf Algorithms \textbackslash{} Hybrid Ciphers}.
}
\label{CM_Hybrid-procedures}
\index{Hybrid procedure}

@@

In this case the bulk data is encrypted using symmetric procedures: The key
used for this is a secret session key\index{Session key} generated by the
sender randomly\footnote{%
   An important part of cryptographically secure techniques is to generate 
   random numbers. Within CT1\index{CrypTool 1} you can check out
   different random number generators using the menu path
   {\bf Indiv. Procedures \textbackslash{} Generate Random Numbers}. 
   Using the menu path {\bf Analysis \textbackslash{} Analyze Randomness}
   you can apply different test methods for random data to binary documents. \\
   CT1 concentrates on cryptographically strong {\bf pseudo} random number
   generators. Only the integrated Secude generator involves a "pure" random source.\\
   JCrypTool\index{JCrypTool} offers {\bf pseudo} random number generators both
   within the Default Perspective in the menu {\bf Algorithms \textbackslash{}
   Random Number Generator} as well as within the Functional Perspective.
}\index{Random}
that is only used for this message.

@@

% --------------------------------------------------------------------------
\newpage

\begin{ctsquote}
    There is an old saying inside the US National Security Agency (NSA):\\
    "Attacks always get better; they never get worse."
\caption[IETF]{IETF\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }

\section[Cryptanalysis and symmetric ciphers for educational purposes]{Cryptanalysis and symmetric ciphers for educational purposes\footnotemark}
\footnotetext{%
A good starting point to learn cryptanalysis is the book from Mark Stamp \cite{cm:Stamp2007}.\\
Very high-level and concentrating on analysing symmetric block ciphers only is the
article from von Bruce Schneier \cite{cm:Schneier2000cm}.\\
Several of the cipher challenges at ``MysteryTwister C3''
(\url{http://www.mysterytwisterc3.org}) are also well suitable for educational purposes.
\index{MTC3}\index{Challenge}\index{Cipher-Challenge}\index{Krypto-Wettbewerb}
}
\index{Cryptanalysis}

@@

\bibitem[Haan2008]{cm:Haan2008} \index{Haan 2008} 
       Kristian Laurent Haan, \\
       {\em Advanced Encryption Standard (AES)},\\
       2008, (German article)\\
       \url{http://www.codeplanet.eu/tutorials/cpp/51-advanced-encryption-standard.html}

@@

\bibitem[Schneier2000]{cm:Schneier2000cm} \index{Schneier 2000} 
       Bruce Schneier, \\
       {\em A Self-Study Course in Block-Cipher Cryptanalysis}, \\
       Cryptologia, v.24, Jan 2000, pp. 18-34\\
       \url{www.schneier.com/paper-self-study.pdf}

### 2012-11-05 script/en/chapters/digitalsignatures.tex
CT Script: Many minor corrections and updates (e.g. 4 levels of titles)
CT script: Many small changes; especially referenced CT2, JCT and Sage where previously only CT1 was mentioned.

\newpage
% --------------------------------------------------------------------------
\hypertarget{Chapter_Hashes-and-Digital-Signatures}{}
\chapter{Hash Functions and Digital Signatures}
\label{Chapter_Hashes-and-Digital-Signatures}
\index{Signature!digital}
(Schneider J. / Esslinger B. / Koy H., June 2002;
Updates: Feb. 2003, June 2005, July 2009)

\vspace{12pt}
\noindent The aim of digital signatures is to guarantee the following two points:
\begin{itemize}
 \item User authenticity: \index{Authenticity!user} \\
      It can be checked whether a message really does
come from a particular person.
 \item Message integrity:  \index{Message integrity} \\
      It can be checked whether the message has been
changed (on route).
\end{itemize}

@@

% --------------------------------------------------------------------------
\vskip + 45pt
\begin{center}
\fbox{\parbox{15cm}{{\em Stanislaw Lem\index{Lem, Stanislaw}\footnotemark:}\newline 
We can make everything out of this world, but we cannot create a world,
where humans in some ten thousand years can think: 'Ok, now it is enough.
Everything should stay like it is. Let's do no changes any more, don't do
inventions any more, because it cannot become better, and if, then we
don't want this.'
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{This was the answer of Stanislaw Lem
to heavy critics at his philosophical main book ~``Summa Technologiae'',
1964, where he thought about the possibility of an evolution creating
artificial intelligence.\\}

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsection{Signing with hash functions}
The signature procedure with hash functions is as follows:\footnote{% 
Compare: \href{http://en.wikipedia.org/wiki/Digital\_signature}
      {\texttt{http://en.wikipedia.org/wiki/Digital\_signature}}.
} \\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

@@

\begin{cryptoprocedure}
\paragraph{Public Key}\strut\\
\begin{tabular}{l@{ }l}
$p$ & prime \\
$q$ & 160-bit prime factor of $p - 1$ \\
$g$ & $ = h^{(p-1)/q}  \mbox{ mod } p$, where $h < p - 1$ and
$h^{(p-1)/q} > 1  \Mod{p}$ \\
$y$ & $\strut \equiv  g^x  \mbox{ mod } p$ 
\end{tabular}

@@

\paragraph{Private Key}\strut\\
\begin{tabular}{l@{ }l}
$x < q$ (a 160-bit number) 
\end{tabular}

\paragraph{Signing}\strut\\
\begin{tabular}{l@{ }l}
$m$ & the message to be signed\\
$k$ & choose at random\index{Random}, less than $q$\\
$r$ & $= (g^k \; \mbox{ mod } p) \mbox{ mod } q$\\
$s$ & $= (k^{-1}(\mbox{SHA-1}(m) + xr)) \mbox{ mod } q$
\end{tabular}

\emph{Remark:}
\begin{itemize}
\item $(s,r)$ is the signature.
\item The security of the signature depends not only on the mathematical
  properties, but also on using a good random source\index{Random}  for $k$.
\item SHA-1 \index{SHA-1} is a 160-bit hash function.
\end{itemize}
\paragraph{Verifying}\strut\\
\begin{tabular}{l@{ }l}
$w$ & $= s^{-1} \;  \mbox{ mod } q$\\
$u_1$ & $= (\mbox{SHA-1}(m)w) \mbox{ mod } q$\\
$u_2$ & $= (rw)  \mbox{ mod } q$\\
$v$ & $= (g^{u_1}y^{u_2}) \mbox{ mod } p)  \mbox{ mod } q$\\

>>>>

\newpage
% --------------------------------------------------------------------------
\hypertarget{Chapter_Hashes-and-Digital-Signatures}{}
\chapter{Hash Functions and Digital Signatures}
\label{Chapter_Hashes-and-Digital-Signatures}
\index{Signature!digital}
(Schneider J. / Esslinger B. / Koy H., June 2002;
Updates: Feb. 2003, June 2005, July 2009)
% --------------------------------------------------------------------------
%\vskip + 30pt
\begin{ctsquote}
We can make everything out of this world, but we cannot create a world,
where humans in some ten thousand years can think: 'Ok, now it is enough.
Everything should stay like it is. Let's do no changes any more, don't do
inventions any more, because it cannot become better, and if, then we
don't want this.'
\caption[Stanislaw Lem]{Stanislaw Lem\footnotemark}\index{Lem, Stanislaw}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{This was the answer of Stanislaw Lem
to heavy critics at his philosophical main book ~``Summa Technologiae'',
1964, where he thought about the possibility of an evolution creating
artificial intelligence.\\}

%\vspace{12pt}
\noindent The aim of digital signatures is to guarantee the following two points:
\begin{itemize}
 \item User authenticity: \index{Authenticity!user} \\
      It can be checked whether a message really does
come from a particular person.
 \item Message integrity:  \index{Message integrity} \\
      It can be checked whether the message has been
changed (on route).
\end{itemize}

@@

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsection{Signing with hash functions}

\begin{ctsquote}
``Manipulation was Sobol's speciality ... the main investigation should brach off and try to discover Sobol's master plan.''
\caption[Daniel Suarez, Daemon]{Daniel Suarez\footnotemark}\index{Suarez, Daniel}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{Daniel Suarez, ``Daemon'', Dutton Adult, 2010,
Chapter 14, ``Meme Payload'', p. 142, Ross.}

The signature procedure with hash functions is as follows:\footnote{% 
Compare: \href{http://en.wikipedia.org/wiki/Digital\_signature}
      {\texttt{http://en.wikipedia.org/wiki/Digital\_signature}}.
} \\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

@@

\begin{cryptoprocedure}
\paragraph*{Public Key}\strut\\
\begin{tabular}{l@{ }l}
$p$ & prime \\
$q$ & 160-bit prime factor of $p - 1$ \\
$g$ & $ = h^{(p-1)/q}  \mbox{ mod } p$, where $h < p - 1$ and
$h^{(p-1)/q} > 1  \Mod{p}$ \\
$y$ & $\strut \equiv  g^x  \mbox{ mod } p$ 
\end{tabular}

@@

\paragraph*{Private Key}\strut\\
\begin{tabular}{l@{ }l}
$x < q$ (a 160-bit number) 
\end{tabular}

\paragraph*{Signing}\strut\\
\begin{tabular}{l@{ }l}
$m$ & the message to be signed\\
$k$ & choose at random\index{Random}, less than $q$\\
$r$ & $= (g^k \; \mbox{ mod } p) \mbox{ mod } q$\\
$s$ & $= (k^{-1}(\mbox{SHA-1}(m) + xr)) \mbox{ mod } q$
\end{tabular}

\emph{Remark:}
\begin{itemize}
\item $(s,r)$ is the signature.
\item The security of the signature depends not only on the mathematical
  properties, but also on using a good random source\index{Random}  for $k$.
\item SHA-1 \index{SHA-1} is a 160-bit hash function.
\end{itemize}
\paragraph*{Verifying}\strut\\
\begin{tabular}{l@{ }l}
$w$ & $= s^{-1} \;  \mbox{ mod } q$\\
$u_1$ & $= (\mbox{SHA-1}(m)w) \mbox{ mod } q$\\
$u_2$ & $= (rw)  \mbox{ mod } q$\\
$v$ & $= (g^{u_1}y^{u_2}) \mbox{ mod } p)  \mbox{ mod } q$\\
